\chapter{Implementation}
\label{chap:implementation}

I have implemented the map-merging algorithm (Chapter~\ref{chap:mergingalgorithm}) in the \gls{ROS} framework (Section~\ref{sec:ros}). This allows the implementation to be used with various readily available \gls{SLAM} algorithms and inter-robot communication solutions.

Point cloud algorithms are implemented with \gls{PCL} (Section~\ref{sec:pcl}). \gls{PCL} is a popular open-source library for manipulating point-cloud data.

\section{Robot Operating System}
\label{sec:ros}

\gls{ROS} is a popular open-source robotics middleware and distribution of community-maintained robotics software. \gls{ROS} provides hardware abstraction, low-level sensors control, implementations of commonly-used robotics algorithms, including several variants of \gls{SLAM}, message-passing between processes, build system, and package management.

The core of the \gls{ROS} is a messaging system, that allows loose coupling between sensors, robotic algorithms and actuators. There are many community-established standard message formats for various sensor types, visualisation, maps, robot state, linear algebra etc., which enable software reuse.

Software in the \gls{ROS} project is organised in a form of community-maintained packages. The first type of packages integrate particular hardware sensors, actuators or entire robot platforms with \gls{ROS} and provide a message interface using standard \gls{ROS} messages. The second type of packages implements robotics algorithms and usually depends only on standard \gls{ROS} messages, making them reusable between robots. The \gls{ROS} project builds and distributes binary packages for Linux.

\section{Point Cloud Library}
\label{sec:pcl}

\gls{PCL} is an open-source library for point cloud processing. It contains routines for feature estimation, surface reconstruction, registration, nearest neighbour search, model fitting, including \gls{RANSAC} algorithm, and segmentation. The library offers serialisation of point clouds to its native, but widely supported \texttt{pcd} file format. \Gls{PCL} is supported in \gls{ROS}, with \gls{PCL} point cloud format being supported for message serialisation into \gls{ROS} messages.

I have used \gls{PCL} to implement the presented map-merging algorithm (Chapter~\ref{chap:mergingalgorithm}) and to read and save \texttt{pcd} files for the command-line tools (Section~\ref{sec:commandline-tools}).

% todo mention dispatch?

\section{\texttt{map\_merge\_3d} package}
\label{sec:ros-package}

The software for this thesis is organised in a \gls{ROS} package \texttt{map\_merge\_3d}. This package contains \gls{ROS} node for online map-merging as well as command-line applications for offline map-merging and visualisation of point cloud files.

The \texttt{map\_merge\_3d} package is distributed within \gls{ROS} starting from \gls{ROS} Melodic release. The package documentation is maintained as wiki text available online, a reproduction of the documentation is attached (Attachment~\ref{chap:map_merge-wiki}).

Care has been taken to integrate the package with the rest of the \gls{ROS} ecosystem. The package is not tied to any particular package for inter-robot communication (Section~\ref{sec:communication}) and it depends on commonly used map representation in \gls{ROS} (Section~\ref{sec:map-representation}).

The \gls{ROS} node supports auto-discovery of the robots (Section~\ref{sec:configuration}) and is designed to partially mitigate the high computing demands of the map-merging (Section~\ref{sec:node-architecture}). The package also contains command-line tools for offline map-merging and visualisation (Section~\ref{sec:commandline-tools}).

\subsection{Communication}
\label{sec:communication}

The \gls{ROS} node is designed to work with any software for inter-robot communication. This architecture allows users to use any kind of communication media and software solution that is available for a given multi-robot system. The package, therefore, does not provide nor expect any particular communication between robots.

To achieve the loose coupling with communication software, the node expects map topics to be available within local \gls{ROS} graph. A user of the package is then expected to configure communication between robots that publishes robot maps within local \gls{ROS} graph. There are several solutions available within \gls{ROS} to achieve this setup.

First of all, \gls{ROS} natively supports sending messages between multiple computers connected to the same network. However, the \gls{ROS} master directory listing (broker) service is not distributed and it needs to run on one of those computers. This creates a single-point-of-failure that is not acceptable for unreliable communication media.

To achieve reliable multi-robot communication \gls{ROS} Multimaster Special Interest Group was formed with the intent to support running multiple \gls{ROS} master broker services. \citet{hernadez2015multi} developed \texttt{multimaster\_fkie} package, that deals with most of the problems of the multi-robot communication. This package can be used together with \texttt{map\_merge\_3d} to run online map-merging for multiple robots.

Users are free to choose any other communication solution meeting their needs for their particular application, \texttt{map\_merge\_3d} does not depend on any particular multi-master setup and can run also under single-master setup, for example in a simulated multi-robot environment.

\subsection{Map representation}
\label{sec:map-representation}

Choosing a correct map representation is important to allow interoperability and re-usability within the \gls{ROS} ecosystem. We need to select a map format that is supported by the most \gls{SLAM} implementations in \gls{ROS} to be able to consume maps from the robots.

Widespread \gls{3D} \gls{SLAM} approaches in \gls{ROS} produce maps as \texttt{sen\-sor\_msgs/\-Point\-Cloud2} messages, which are now de-facto standard for representing \gls{3D} maps in \gls{ROS}. This representation is suitable for the presented map-merging algorithm (Chapter~\ref{chap:mergingalgorithm}).

Unfortunately, there is no standard message format in \gls{ROS} to represent maps as a pose graph of point clouds, which would allow implementing map-merging on point cloud pose graph discussed in Section~\ref{sec:map-merging-on-point-clouds}, which allows repairing mapping errors.

The map-merging node, therefore, uses monolithic maps represented in \texttt{sen\-sor\_msgs/\-Point\-Cloud2} messages, which always encode the whole map, to allow the highest degree of interoperability with existing \gls{ROS} packages.

Apart from not being able to correct mapping errors, a monolithic map representation also causes unnecessary transfers of already-explored space, that slows map update rate for large-scale maps. A standardised pose graph map representation in \gls{ROS} might help to solve both issues.

% mention rtabmap, google cartographer?

\subsection{Configuration}
\label{sec:configuration}

The \gls{ROS} node supports merging maps from an arbitrary number of robots. To make configuration of the node easy, maps for map-merging are auto-discovered by the \gls{ROS} node.

The node periodically scans the \gls{ROS} graph to discover map topics, which are then subscribed and passed to the map-merging procedure. This mechanism also enables robots to be added to the multi-robot system later, for example, to tackle a high demand for the service.

Full documentation of this mechanism is described in Attachment~\ref{chap:map_merge-wiki}.

\subsection{ROS node architecture}
\label{sec:node-architecture}

Estimating transformations between maps with the presented map-merging algorithm (Chapter~\ref{chap:mergingalgorithm}) is computationally intensive and can take a long time, especially when merging many large-scale maps.

To achieve high update rates of the merged map, the \gls{ROS} node uses an asynchronous architecture. Because the map origins do not change, we don't need to re-estimate transformations between maps every time we are updating the merged map, we can use the previous estimates to update the map.

The node is designed to run periodically asynchronous map composition and transformations estimation tasks, which are independent, at user-settable frequencies. The map composition uses already estimated transformations between maps, transforms the input maps and concatenates the maps to produce the merged map. This process is fast and allows the node to achieve high update rates of the merged map to quickly incorporate newly-discovered areas by the robots.

The transformations estimation task uses a computationally demanding algorithm (Chapter~\ref{chap:mergingalgorithm}) to find the transformations between maps with enough overlapping space. This task can run with a lower frequency to save computational resources.

\subsection{Offline map-merging and visualisation}
\label{sec:commandline-tools}

The \gls{ROS} package contains also two command-line applications besides the main \gls{ROS} node. The applications work with point clouds saved in \texttt{pcd} files, a point cloud file format of the \gls{PCL} library.

The first application, \texttt{map\_merge\_tool}, is designed for the offline map-merging. It accepts $n$ point cloud files and produces a single merged map saved to a file. It uses the same algorithm for map-merging as the \gls{ROS} node. This application may be used to create a complete global map of a large-scale environment for deployment of a robotic system after a multi-session mapping of the environment with a single robot.

The second application, \texttt{registration\_visualisation}, accepts two point clouds and visualises pair-wise transformation estimation algorithm as described in Chapter~\ref{chap:mergingalgorithm}. The tool is mostly controlled by the command-line arguments, but the visualisation window is graphical. Each step of the estimation is visualised in \gls{3D}, the user is able to freely navigate the point clouds or save the visualisation as an image. This application is intended for estimation parameter tuning and learning about the map-merging process.

% todo screenshot
